---
title: "Effects of Sleep Loss"
author: "Loecher"
date: '2022-08-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(lubridate)
library(readxl)
library(emmeans)
```

Sleep loss leads to the withdrawal of human helping across individuals, groups, and large- scale societies

Eti Ben Simon1, Raphael Vallat1, Aubrey Rossi1, Matthew P. Walker

```{r}

pbio <- read_excel("journal.pbio.3001733.s008.xlsx")
pbio$day_of_year = yday(pbio$donation_date) 
pbio_all <- read_excel("journal.pbio.3001733.s008.xlsx", sheet = "S3_Data_FigS2")
length(unique(pbio_all$donation_date))
```



## Recreating Figure 3A


```{r}
hist(log(pbio_all$daily_donations_mean))
mean(pbio_all$daily_donations_mean)
```

It almost looks as if the authors displayed the median

```{r}
median(pbio_all$daily_donations_mean)
x = pbio_all$daily_donations_mean
SE = mad(pbio_all$daily_donations_mean)/sqrt(nrow(pbio_all))*1.96
print(SE)
```


```{r}
#exp(mean(log(pbio_all$daily_donations_mean)))
```

```{python}
import seaborn as sns
import numpy as np
df = r.pbio_all

sns.kdeplot(data=df, x=np.log(df.daily_donations_mean), fill=True,bw_adjust=1.5);

```

```{r}

```

## Recreating Figure 3B

```{r}

lm3 <- glm(log(daily_donations_mean) ~ Type_of_week+Time_of_Day+ Weekend_day+ YEAR+MONTH+DAY_OF_WEEK+Holiday, family="gaussian", data=pbio)


summary(lm3)
```

Why not a vanilla linear model ?

```{r}
lm3 <- lm(log(daily_donations_mean) ~ Type_of_week+Time_of_Day+ Weekend_day+ YEAR+MONTH+DAY_OF_WEEK+Holiday,  data=pbio)
summary(lm3)
```



```{r}
emmeans(lm3, ~ Type_of_week)
plot(emmeans(lm3, ~ Type_of_week))
```
Do we not have to code up month and  DAY_OF_WEEK as factors?

```{r}
pbio_dummy = pbio
pbio_dummy$MONTH = as.factor(pbio_dummy$MONTH)
pbio_dummy$DAY_OF_WEEK = as.factor(pbio_dummy$DAY_OF_WEEK)

lm_dummy <- lm(log(daily_donations_mean) ~ Type_of_week+Time_of_Day+ Weekend_day+ YEAR+MONTH+DAY_OF_WEEK+Holiday,  data=pbio_dummy)
summary(lm_dummy)
```

## Constructing our own data

```{r}
library(timeDate);#library(chron)
#As of June 2021, there are eleven annual federal holidays in the United States, and one additional quadrennial holiday (Inauguration Day)
#There must be more school holidays, possibly different from state to state

hlist <- c("USChristmasDay","USGoodFriday","USIndependenceDay","USLaborDay",
           "USNewYearsDay","USThanksgivingDay")        
holidays2000_2016  <- as.Date(timeDate::holiday(2000:2016,hlist))

#is.holiday(as.Date("2013-11-28"),myholidays)
```


```{r}
pbio_all$MONTH = lubridate::month(pbio_all$donation_date)
pbio_all$YEAR = lubridate::year(pbio_all$donation_date)
pbio_all$DAY_OF_WEEK = lubridate::wday(pbio_all$donation_date)
#try lubridate::wday(as.Date("2023-02-19"))
pbio_all$Weekend_day = (pbio_all$DAY_OF_WEEK %in% c(1,7))#Sunday=1, Saturday=7!
pbio_all$Holiday = as.Date(pbio_all$donation_date) %in% holidays2000_2016
pbio_all$day_of_year = lubridate::yday(pbio_all$donation_date) 
#make lookup table:
Type_of_week = as.data.frame(unique(pbio_dummy[,c("donation_date", "Type_of_week", "Holiday")]))
rownames(Type_of_week) = Type_of_week$donation_date
pbio_all$Type_of_week = Type_of_week[as.character(pbio_all$donation_date), "Type_of_week"]
sum(is.na(pbio_all$Type_of_week))
pbio_all$Type_of_week[is.na(pbio_all$Type_of_week)] = "Regular"

pbio_all$MONTH = as.factor(pbio_all$MONTH)
pbio_all$DAY_OF_WEEK = as.factor(pbio_all$DAY_OF_WEEK)
#mean(pbio_dummy$donation_date %in% pbio_all$donation_date)
#table(table(pbio_dummy$donation_date))
#table(table(pbio_all$donation_date))

#tmp = merge(pbio_all, pbio_dummy[,c("donation_date", "Type_of_week")], by = "donation_date", all.x = TRUE)

lm_dummy <- lm(log(daily_donations_mean) ~ Type_of_week+ Weekend_day+ YEAR+MONTH+DAY_OF_WEEK+Holiday,  data=pbio_all)#+Time_of_Day
summary(lm_dummy)

```

## GAM models

```{r}
library(mgcv)
gam_dummy <- gam(log(daily_donations_mean) ~ Type_of_week+ Weekend_day+ s(YEAR, k=3) +DAY_OF_WEEK+Holiday + s(day_of_year,k=8), family=gaussian(),  data=pbio_all)#+Time_of_Day +MONTH
summary(gam_dummy)
plot(gam_dummy)
```


## Old Code

```{r}
library(dplyr)

pbioDonations <-  pbio %>% group_by(donation_date) %>% summarise(daily_donations_sum = sum(daily_donations_mean), Type_of_week = unique(Type_of_week)) %>% 
  group_by(Type_of_week) %>% summarise(daily_donations_avg = mean(daily_donations_sum), daily_donations_sd = sd(daily_donations_sum), cts = n())
```


```{r}
hist(pbio$daily_donations_mean)
```

```{r}
n = 2:20
plot(n, (n+2)/n/(n+1), type = "l");grid()
lines(n, 1/(n-1), type = "l", col=2)
lines(n, 1/(n), type = "l", col=3)

```

